# short

[![Build Status](https://travis-ci.org/haplo/short.svg?branch=master)](https://travis-ci.org/halpo/pivot)
<!--[![CRAN\_Status\_Badge](http://www.r-pkg.org/badges/version/pivot)](https://cran.r-project.org/package=short)-->

```{r setup, echo=FALSE, results='hide'}
library(short)
```

# Overview

The `short` package gives several shortcuts to common functionality.

# `using` - easy loading packages

The `using` function is a simpler way of specifying multiple packages to 
load.  It additionally will attempt to download and install the packages
if not already installed.


# `table_1` - Table 1 Demographics and Summaries

The `table_1` function can be used to generate a demographics and summary
of variables; i.e. the veritable "Table 1" for publications.
For each variable given, `table_1` calls `summarise_table_1` which 
will summarize the variable for each value of a key, differentiating 
variable, as well as overall.  For character and factor variables, 
the counts and percentages of each level is given.
For logical variables, the count of positives and the percentage is 
given as the value.  Numeric Variables are summarized by the minimum, 
median, mean, maximum, and standard deviation.

Additional methods for table_1 can be added by adding methods to 
`summarise_table_1`

### Example

```{r table_1_example, df_print='kable'}
library(tidyverse)
mtcars %>% 
    mutate( gear = ordered(paste(gear, 'Gears'))
          , cyl  = ordered(paste(cyl, 'Cylinders'))
          ) %>% 
    table_1( cyl
           , 'Automatic Transmission' = am == 0
           , 'Number of gears' = gear
           , Horsepower = hp
           )
```


# `with_margins` - add marginals to and computation.

The `with_margins` function converts another function to operate over 
all possible subsets of groups, including no groups.

### Example

```{r with_margins_example}
mtcars %>% 
    mutate( gear = ordered(paste(gear, 'Gears'))
          , cyl  = ordered(paste(cyl, 'Cylinders'))
          ) %>%
    group_by(gear, cyl) %>% 
    with_margins(summarise_at)( vars(mpg, disp, hp, wt)
                              , funs(min, mean, max)
                              )
```


# `spread_each` - Spread Multiple Values

The `spread_each` function works similar to the `tidyr::spread` function 
but works for spreading out multiple values.

### Example

```{r spread_each_example}
mtcars %>% 
    count(gear, cyl) %>% 
    group_by(gear) %>%
    mutate(pct = n/sum(n)) %>% 
    spread_each(cyl, n, pct)
```


# `.T` - Simple text specification

Similar to lazy loading of the [`tidyverse`](https://tidyverse.org), 
the `.T` function provides
a simple way to designate arguments should be interpreted as text.
The limiting factor is that the text must be syntactically correct, 
however it come in very helpful.

### Example

```{r T_example}
.T(first, second, 'a third argument with spaces')
```


# Inline functions

The `short` package also provides many inline shortcuts.

## Text concatenations

* `%\\n%` - concatenate with a line break
* `%<<%` - concatenate with a space
* `%<<<%` - concatenate without space.

## Pattern Matching

* `%~%` - Matches the pattern on the left hand side.
* `%!~%` - Does not match the pattern  on the left hand side.

## Other

* `%or%` - Similar to the `rlang` or `purrr` function `%||%` but checks
    on `length` rather than `is.null` so will give the right hand 
    side for length 0 vectors as well.
* `%inherits%` - inline version of `inherits`
* `%is a%` - alternate inline version of `inherits`


# String Filtering

There are two convenience function for filtering character vectors, 
`sift` and the filter out version `seive`.

```{r}
# limit to Mercs
mtcars %>% rownames() %>% sift("^Merc")
# Filter out everything starting with M
mtcars %>% rownames() %>% sieve("^M")
```
